### 2017-05-23 R/NHTS.Rmd
Updates to TravelSurvey.Rmd and NHTS.Rmd

# Integrated Transport and Health Impacts Model

## R Package _ITHIM_

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
options(repos='http://cran.rstudio.com/')
#install.packages(c("ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
library("ggplot2")
library("devtools")
library("reshape2")
library("plyr")
library("dplyr")
library("TravelSurvey")
```

```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
document("~/ITHIM/")
install("~/ITHIM/")
#install_github("syounkin/ITHIM", ref="devel")
install_github("smbache/magrittr", ref="master")

library("magrittr")
library("ITHIM")

fig.width <- 6 # inches
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

## HHS Metro Region Parameter Sets

First we use NHTS to estimate mean walking and cycling times.

### NHTS

@@

```{r parsely, eval = TRUE, echo = FALSE}
convertToHHS <- function(states){
  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS01","HHS01","HHS01","HHS01","HHS01","HHS01","HHS02","HHS02","HHS03","HHS03","HHS03","HHS03","HHS03","HHS03","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS05","HHS05","HHS05","HHS05","HHS05","HHS05","HHS06","HHS06","HHS06","HHS06","HHS06","HHS07","HHS07","HHS07","HHS07","HHS08","HHS08","HHS08","HHS08","HHS08","HHS08","HHS09","HHS09","HHS09","HHS09","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
  StateToHHS <- HHSRegions[,3]
  names(StateToHHS) <- HHSRegions[,1]
  return(StateToHHS[states])
}

countPeople <- function(data){
 nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
}

tripTime <- function(data, mode = "walk"){
 return(sum(subset(data, AT == mode)$TRVL_MIN, na.rm = TRUE))
}

reformatData <- function(data){
foo <- matrix(unlist(data), nrow = 7, ncol = 2)
foo <- rbind( c(0, 0), foo)
dimnames(foo) <- list(paste0("ageClass",1:8), c("F","M"))
return(foo[,c("M","F")])
}

#### Read Two of the Four NHTS Data Tables

```{r NHTS, eval = compute, echo = FALSE}
DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")

@@

NHTS.person <- with(PERV2PUB,{
                    data.frame(HOUSEID = HOUSEID,
                               PERSONID = PERSONID,
                               SEX = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA))),
                               AGE = as.factor(ITHIM:::convertToAgeClass(R_AGE)),
                               MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
                               HHSTATE = as.character(HHSTATE),
                               HHS = as.character(convertToHHS(as.character(HHSTATE))),
                               MSA_ID = HHC_MSA, stringsAsFactors = FALSE)
})

@@

#### Create TravelSurvey Object

```{r TravelSurvey, eval = compute, fig.width = 10, echo = FALSE}

@@

```{r TravelSurvey2, eval = compute, fig.width = 10, echo = FALSE}
NHTS.TravelSurvey <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")

NHTS.df <- dplyr::left_join(NHTS.TravelSurvey@person, NHTS.TravelSurvey@trip, by = c("houseID","subjectID"))

NHTS.df <- within(NHTS.df, duration <- ifelse(is.na(duration), 0, duration))

```{r msaMatrix, eval = FALSE, fig.width = 10, echo = FALSE}
msaMatrix <- matrix(c(5602,4472,1922,7320,3120,6200,3362,7362,5720,4992,8872,
8280,7240,0640,6922,5960,3600,0520,8960,3480,1602,6840,
1280,1520,5360,6640,1122,5082,6162,2162,4920,7602,6480,
5120,1692,7040,7160,6442,4120,4520,6280,2082,1642,3760,
3280,1840,5880,3000,5560,
"NY/NJ Metro","Los Angeles-Riverside-Orange County","Dallas-Fort Worth","San Diego","Greensboro-Winston Salem", "Phoenix/Mesa","Houston-Galveston-Brazoria", "San Francisco-Oakland-San Jose", "Norfolk-Virginia Beach", "Miami-Ft. Lauderdale", "Washington-Baltimore", rep("no label yet", 38)), ncol = 2)
```

```{r foob, eval = FALSE, echo = FALSE}
#NHTS.person <- subset( NHTS.person, MSA == TRUE )
#NHTS.trips <- subset(NHTS.trips, !is.na(TRVL_MIN))

#saveRDS(NHTS.person, file = "./data/NHTS.person.rds")
#saveRDS(NHTS.trips, file = "./data/NHTS.trips.rds")

#NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"), all.x = TRUE)
#NHTS <- within(NHTS, TRVL_MIN <- ifelse(is.na(TRVL_MIN), 0, TRVL_MIN))
#saveRDS(NHTS, file = "./data/NHTS.rds")
```

>>>>

# U.S. National Household Travel Surevy

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
options(repos='http://cran.rstudio.com/')
#install.packages(c("ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
library("tidyverse")
library("devtools")
#library("reshape2")
#library("plyr")
#library("dplyr")
library("TravelSurvey")
library("ITHIM")
fig.width <- 6 # inches
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

@@

### Read Two of the Four NHTS Data Tables

```{r NHTS, eval = compute, echo = TRUE}
DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")

@@

NHTS.person <- with(PERV2PUB,{
                    data.frame(HOUSEID = HOUSEID,
                               PERSONID = PERSONID,
                               SEX = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA))),
                               AGE = as.factor(ITHIM:::convertToAgeClass(R_AGE)),
                               MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
                               HHSTATE = as.character(HHSTATE),
                               # HHS = as.character(convertToHHS(as.character(HHSTATE))),
                               MSA_ID = HHC_MSA, stringsAsFactors = FALSE)
})

@@

### Create TravelSurvey Object

```{r TravelSurvey, eval = compute, fig.width = 10, echo = TRUE}

@@

### 2017-05-23 R/TravelSurvey.Rmd
Updates to TravelSurvey.Rmd and NHTS.Rmd

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
options(repos='http://cran.rstudio.com/')
library("tidyverse")
library("ITHIM")
library("TravelSurvey")
```

@@

```{r TravelSurvey, eval = compute, fig.width = 10, echo = FALSE}
NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
NHTS.df <- dplyr::left_join(NHTS.ts@person, NHTS.ts@trip, by = c("houseID","subjectID"))
NHTS.df <- within(NHTS.df, {
    duration <- ifelse(is.na(duration), 0, duration)
    mode <- factor(mode, levels = c("walk","cycle","other"))
}
)

@@

```{r plot1, eval = TRUE, fig.width = 20, echo = FALSE}
total <- NHTS.df %>% group_by(location) %>% summarise(total = sum(duration) )
byMode <- NHTS.df %>% group_by(location, mode) %>% summarise(byMode = sum(duration) )
NHTS.table1 <- dplyr::full_join(byMode, total, by = "location") %>% mutate(propMode = byMode/total) %>% dplyr::filter(mode %in% c("walk","cycle"))

p <- ggplot(data = NHTS.table1, aes(x = location, y = propMode, fill = mode))
p + geom_bar(stat="identity") + theme_bw()
```

>>>>

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
library("tidyverse")
library("ITHIM")
library("TravelSurvey")
```

@@

```{r TravelSurvey, eval = compute, fig.width = 10, echo = FALSE}
NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
NHTS.df <- dplyr::left_join(NHTS.ts@person, NHTS.ts@trip, by = c("houseID","subjectID"))
NHTS.df <- within(NHTS.df, {
    duration <- ifelse(is.na(duration), 0, duration)
    mode <- factor(mode, levels = c("walk","cycle","other"))
})
```

```{r plot1, eval = TRUE, fig.width = 20, echo = FALSE}
total <- NHTS.df %>% group_by(location) %>% summarise(mean = mean(duration), total = sum(duration), sdTotal = sd(duration), nTotal = n())

byMode <- NHTS.df %>% group_by(location, mode) %>% summarise(byMode = mean(duration), sdMode = sd(duration), nMode=n() )
NHTS.table1 <- dplyr::full_join(byMode, total, by = "location") %>% mutate(propMode = byMode/total) %>% dplyr::filter(mode %in% c("walk","cycle")) %>% mutate(ciModeLeft = byMode - 1.96*sdMode/sqrt(nMode), ciModeRight = byMode + 1.96*sdMode/sqrt(nMode))

p <- ggplot(data = NHTS.table1, aes(x = location, y = propMode, fill = mode))
p + geom_bar(stat="identity") + theme_bw()

NHTS.table2 <- NHTS.table1 %>% select(location, mode, ciModeLeft, ciModeRight) %>% gather(ciMode, value, 3:4)
NHTS.table2 <- within(NHTS.table2, {
    location <- factor(location, levels=unique(location[order(value)]))
    })

p <- ggplot(NHTS.table2, aes(x = location, y = value, colour = mode))
p + geom_line(aes(group = ciMode))# + facet_grid(mode ~ . )

@@

### 2017-05-23 R/appendix.Rmd
Updates to TravelSurvey.Rmd and NHTS.Rmd

### Appendix

>>>>

### Appendix
### Appendix

```{r msaMatrix, eval = FALSE, fig.width = 10, echo = FALSE}
msaMatrix <- matrix(c(5602,4472,1922,7320,3120,6200,3362,7362,5720,4992,8872,
8280,7240,0640,6922,5960,3600,0520,8960,3480,1602,6840,
1280,1520,5360,6640,1122,5082,6162,2162,4920,7602,6480,
5120,1692,7040,7160,6442,4120,4520,6280,2082,1642,3760,
3280,1840,5880,3000,5560,
"NY/NJ Metro","Los Angeles-Riverside-Orange County","Dallas-Fort Worth","San Diego","Greensboro-Winston Salem", "Phoenix/Mesa","Houston-Galveston-Brazoria", "San Francisco-Oakland-San Jose", "Norfolk-Virginia Beach", "Miami-Ft. Lauderdale", "Washington-Baltimore", rep("no label yet", 38)), ncol = 2)
```

```{r foob, eval = FALSE, echo = FALSE}
#NHTS.person <- subset( NHTS.person, MSA == TRUE )
#NHTS.trips <- subset(NHTS.trips, !is.na(TRVL_MIN))

#saveRDS(NHTS.person, file = "./data/NHTS.person.rds")
#saveRDS(NHTS.trips, file = "./data/NHTS.trips.rds")

#NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"), all.x = TRUE)
#NHTS <- within(NHTS, TRVL_MIN <- ifelse(is.na(TRVL_MIN), 0, TRVL_MIN))
#saveRDS(NHTS, file = "./data/NHTS.rds")
```

```{r parsely, eval = FALSE, echo = FALSE}
convertToHHS <- function(states){
  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS01","HHS01","HHS01","HHS01","HHS01","HHS01","HHS02","HHS02","HHS03","HHS03","HHS03","HHS03","HHS03","HHS03","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS05","HHS05","HHS05","HHS05","HHS05","HHS05","HHS06","HHS06","HHS06","HHS06","HHS06","HHS07","HHS07","HHS07","HHS07","HHS08","HHS08","HHS08","HHS08","HHS08","HHS08","HHS09","HHS09","HHS09","HHS09","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
  StateToHHS <- HHSRegions[,3]
  names(StateToHHS) <- HHSRegions[,1]
  return(StateToHHS[states])
}

countPeople <- function(data){
 nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
}

tripTime <- function(data, mode = "walk"){
 return(sum(subset(data, AT == mode)$TRVL_MIN, na.rm = TRUE))
}

reformatData <- function(data){
foo <- matrix(unlist(data), nrow = 7, ncol = 2)
foo <- rbind( c(0, 0), foo)
dimnames(foo) <- list(paste0("ageClass",1:8), c("F","M"))
return(foo[,c("M","F")])
}
